\chapter{Report Body}
The central part of the report usually consists of three or four chapters detailing the technical work undertaken during the project. {\bf{\textcolor{red}{The structure of these chapters is highly project dependent}}}. They can reflect the chronological development of the project, e.g. design, implementation, experimentation, optimisation, evaluation, etc (although this is not always the best approach). However you choose to structure this part of the report, you should make it clear how you arrived at your chosen approach in preference to other alternatives. In terms of the software that you produce, you should describe and justify the design of your programs at some high level, e.g. using OMT, Z, VDL, etc., and you should document any interesting problems with, or features of, your implementation. Integration and testing are also important to discuss in some cases. You may include fragments of your source code in the main body of the report to illustrate points; the full source code is included in an appendix to your written report.

\section{Section Heading}

\subsection{Subsection Heading}


A regular wall mountable phone charger usually provides around 1 Ampere of current at 5 Volts. A simple solution for a power supply is to take a phone charger with a detachable inexpensive \emph{USB micro-B} cable, remove the micro-B connector and solder breadboard pins onto the \emph{Gnd} (Ground) and \emph{Vcc} (+5V) pins. As such, an important consideration for hardware choice will be compatibility with a +5V power supply.
