\chapter{Legal, Social, Ethical and Professional Issues and application of the BCS Code of Conduct}

\section{Legal, Social, Ethical and Professional Issues}

Taken into account the nature of this project and the way it has been conducted,
Legal, Social, Ethical and Professional Issues were unlikley to arrise. That being
said, it is important to go over each field separatley and show that no issues pertaining
to that field have appeard.

\begin{itemize}
  \item \emph{Legal issues: } The project did not pose any legal issues. There was no need
  to contact third parties, gain access to sensitive or classified information or any other
  activity which chould have created legal issues.
  \item \emph{Social or Ethical issues: } The project did not pose any social or ethical issues.
  There was no need to gather personal data, conduct surveys, create control groups or anything
  similar which could have affected the project and create social or ethical issues.
  \item \emph{Professional issues: } one professional issue presented itself early on in the
  project. As the project had a significant hardware component, the need to purchase the hardware
  became aparrent. The issue at hand was to determine who would fund the hardware needed for the
  project. After careful consideration, the decision was made to fund the project using personal funds
  so as to not have to be dependant on any other body or entity.
\end{itemize}


\section{Application of the BCS Code of Conduct}
The British Computer Society Code of Conduct \cite{bcs} applies to the project and, as such, the
project abides by this code of conduct. The main section where it is visible that this report aims to
abide by the BCS Code of Conduct is the ``Duty to the Profession'' section, point ``f'', as  well as
the ``Duty to the Profession'' interpretation. They state that a fellow abiding by the BCS Code
of Conduct shall ``encourage and support fellow members in their professional development'' and ``share
knowledge and understanding of IT and support inclusion of every sector of society.''. One of the main
goals of the project is to make computer architecture and design easier to approach from any
potential background and learn more about how computers work at a very low level. This is a reflection
of the directives cited above.
